### 2017-09-26 response/response-jde-1.tex
Add response to R1's question on abstinence
Question was whether abstinence could be recorded as a contraceptive
strategy even if it was caused by other factors, such as stress.
Added table to contraceptive effectiveness file that subdivides
contraceptive use so abstinence only is a separate category,
and updated Makefile and main paper with this new appendix table.

Fixed the definitions in exclude abstinence file.

Wrote initial response based on new results to argue that if
stress could cause this effect we should see something similar
with illness but there is no evidence of that.

\documentclass[letterpaper,12pt]{article}

% Xetex preamble
\usepackage{fontspec}
\setromanfont[Ligatures=TeX]{TeX Gyre Pagella}
\usepackage{unicode-math}
\setmathfont{TeX Gyre Pagella Math}

\usepackage[longnamesfirst]{natbib}
%\usepackage[tabhead,nolists,tablesfirst]{endfloat}
\usepackage[flushleft]{threeparttable}
\usepackage{booktabs}
\usepackage{rotating}  
\usepackage{amsmath}
\usepackage{caption} 
\usepackage{dcolumn} 
\usepackage{setspace}
%\usepackage{flafter} 
%\usepackage{longtable}
%\usepackage[pdftex]{graphicx}
\usepackage[xetex,colorlinks=true,linkcolor=black,citecolor=black]{hyperref} 
\usepackage[margin=1.0in]{geometry} 
\usepackage{multirow}

%opening
\title{} \author{}

\doublespacing

\begin{document}

\begin{center} \textbf{\large Income Shocks, Contraceptive Use, and
Timing of Fertility} \end{center}

\begin{center} Response to Referee Comments \end{center}

\noindent Please find attached the revised version of our paper,
``Income Shocks, Contraceptive Use, and Timing of Fertility.''
We would like to thank the two referees for the very useful comments and
suggestions,
which we believe has improved the paper.

[Additional discussion of overall things we have changed---if any]

We rewrote the code for cleaning and analyzing the data from the ground up 
to make it easier to follow and eventually publish on GitHub.
This rewrite slightly lowered the sample used by three observations.
Including all married/partnered women up to age 45, rather than exclude
women younger than 18, the final sample size is, however, the same as 
before and the results are essentially unchanged.

We list below our responses to the individual comments and suggestions.

\newpage

\section*{Reviewer 1}

% This interesting paper analyzes the relationship between income shocks
% (as proxied by self-reported crop loss) and marital fertility, using a
% small Tanzanian panel survey from the early 1990s. A key feature of the
% paper is that the authors are able to observe use of both traditional
% and modern birth control techniques. In regression specifications with
% and without community or woman fixed effects, the authors find that crop
% loss is associated with decreases in the probability of conceiving, with
% a corresponding increase in use of traditional birth control techniques
% but not modern contraceptives.

The paper could benefit from focusing its motivation somewhat more. An
important point (that the authors do not emphasize enough) is that use
of modern contraception in the sample is extremely low, so the lack of a
relationship between shocks and use of modern methods is not surprising.
The more noteworthy result is that use of traditional methods responds
to shocks. In this sense, the paper makes an important contribution to
understanding how populations control fertility before widespread
adoption of modern contraception. For instance, it is well known that
marital fertility in Europe declined largely before modern contraception
became available, but the methods couples used to achieve this reduction
are not well documented (see, for example, Guinnane JEL 2011). The paper
would be even more compelling if the authors focused the paper's
motivation on the questions of whether and how couples modify fertility
to respond to changing economic conditions before the advent of
modern contraception. That motivation would also help justify their
choice of dataset, which otherwise seems somewhat small and out of date
relative to other available panel datasets from developing countries.

[RESPONSE:]

[Also include Cinnirella since that covers a similar topic]

\noindent \textbf{Specific comments:}

\begin{enumerate}

\item The authors interpret their results as postponement, implying they
represent a change in the timing of fertility but not its lifetime
cumulation. That is likely to be the correct interpretation, but can
they really tell? It seems to me that the current fixed effect model
does not tell us. Perhaps they could include 1-2 lags in the crop shock
variable to shed some light on it.

[RESPONSE:]

It is correct that the current fixed effects model cannot tell
us whether we observe postponement of fertility or a permanent
reduction.
We focus on postponement for two reasons.

First, active postponement is a necessary---but not sufficient---condition
for permanent fertility reduction.
If we cannot show a conscious decision to delay fertility, it is 
hard to argue that we can ever find evidence of conscious permanent 
reduction in fertility.

Second, it is not clear that shocks---in and off themselves---should
substantially affect completed fertility. 
We discuss evidence from Guatemala that the overall \emph{risk} of 
experiencing a shock is likely to affect fertility, but that only
shocks that occur late in a woman's reproductive life cycle is
likely to affect completed fertility (because the opportunities to
make up for the ``missed'' birth are fewer and because the marginal
benefit of a child is smaller the more children you already have and
the older you are).
Hence, an important distinction is that long term fertility is likely
affected by the overall \emph{risk}, whereas the exact timing of
fertility is driven by a combination of risk and shocks (probably
mostly shocks).

The longitudinal nature of the data is important for our ability
to control for unobservable village, household, and individual 
characteristics, but the trade-off is that the data only cover 
32 months and that is too short to analyze long-term fertility
outcomes.

Claus: run main regression with longer lags to check if any effects
on pregnancy and birth

Claus: write something on Guatemala paper and how to think about how
long it takes to find an effect / recover from a shock

There are two competing factors here.
First, a negative shocks will make people more likely to postpone 
their pregnancy/birth because of the factors discussed in the
paper.
Second, longer spacing between births is costly because economies
of scale are harder to realize and because it means that the mother
will be out of the labor force for longer.

This might explain why there are negative effects for both shock
and lagged shock, but the interaction between the two is positive.

One problem with the regression is that we lose 1/4 of observations
each time we include an additional lag.

I have also added an interaction between crop loss for different periods
with the idea that there is a cost to continually postponing births. If
that holds, the effect of both shocks on their own should be same and
the interaction the opposite. Broadly speaking that appears to hold. The
upshot–as far as I can tell–is that what we observe is postponement and
not an overall limitation in fertility. This makes sense given my
results for Guatemala, where long-term risk affects the desire fertility
*level* whereas shocks experienced affect the *timing* of fertility. I
will try to write something more up on this.

\item The authors argue that the increase in modern contraceptive use is
intentional, but the evidence for this interpretation is weaker than
they would like. As they do note, while the survey asks about using
abstinence specifically to prevent births, there is much scope for
misreporting, and many women may claim that reductions in coital
frequency were related to fertility planning even if they were not. As
their main evidence against this concern, the authors show that crop
loss has small effects on labor supply. But this finding is only
relevant for the very specific theory that all unintentional abstinence
is related to a work-sex tradeoff. Couples may have less sex during hard
times because they are stressed, for example.

[RESPONSE:]

[Need to rewrite the section to cover both work and illness]
[A-12 for any, except abstinence only, is very close to statistically significant]

We assume that the reference to ``modern'' is a typo and that the
first sentence is meant to refer to traditional contraceptives.
To be clear, we argue that both modern and traditional contraceptive
use is intentional.
It is correct that we cannot directly rule out that some women will 
report abstinence as a contraceptive strategy even if the abstinence 
arise from something other than a desire to limit the risk of pregnancy.
It is important, however, to note that the question was specifically
about contraception.

\begin{quote}
Some couples use contraception methods to avoid pregnancy or to
space births. Are you currently using a method of contraception?
(For example, the pill, the IUD, condom, withdrawal, rhythm,
abstinence.)
\end{quote}

Although we cannot measure directly measure stress, we do have information 
on illnesses and their durations.
If stress is an important factor in reporting abstinence, we should
expect the same to be the case with illness, especially those that 
prevent a woman from working.
Hence, although we did not explicitly cover this possibility in the original 
version, coital frequency may also decrease because of illness.
% We would expect illness to go up both because of the direct effect
% on the household and from increased stress.
We show the effect of crop loss on illness in Table 5.
Crop loss is  not associated with substantial enough changes in BMI or 
self-reported illness for the women in the sample to explain our fertility results.
Furthermore, new regression we ran show that there is only a very weak 
positive association between having been ill for two or more days in the last 
6 months and the likelihood of reporting using abstinence as a contraceptive 
strategy.
This relationship becomes only marginally stronger, and never anywhere near
statistically significant, if defining illness as not being able to
perform usual activities for 7 days or more.
Going above 7 days makes the relationship negative, most likely because
very few women report being unable to perform regular activities for that
long.

To further investigate the role of abstinence we added a new
table to the Appendix, which further subdivides the results in Table 6.
Table A-13 [TK] show, not surprisingly, that abstinence is the most 
effective traditional contraception, although only in combination 
with a crop loss as in the table in the main text.
A major caveat is that only 19 women report using abstinence only
across the three waves used for these regressions.
Modern contraceptive also are mainly effective in combination with
a crop loss, although the direct effect of using modern contraceptives
is not statistically significant.

[TK summing up]

% Reduction in pregnancy is 0.3 for abstinence with crop loss
% Reduction in pregnancy is 0.2 for modern contraceptives with crop loss

\item The authors analyze the fertility of women who were married at the
start of the sample period. As a separate analysis, it would be
interesting to know whether marriage hazards change with crop loss in
the parents' home, although perhaps the authors do not have the sample
size to study this question.

[RESPONSE:] That is, indeed, a very interesting question, but we have 
decided not to pursue it at the moment for three reasons.
First, there are already two papers that examine this question, 
albeit for other countries \citet{mbiti2008,Hoogeveen2011}.
Second, as those two papers show, the length of our paper would have
to increase substantially to do this question justice, and the Editor
has asked us to not increase the length of the paper.
Finally, the sample size is, indeed, on the small size to really get 
at this question.

We may return to this question in a separate paper.
A problem with the current literature on timing of marriage
is that it tend to focus on only the female side of the marriage market.
A more satisfactory model will have to consistently explain both sides 
of the market.
Hence, results that show, for example, that women are less likely to
marry after a crop loss would have to show the same for men.

\item The survey measures crop loss continuously, and the indicator for
crop loss greater than 200 TZS is the authors' creation. They show in
the appendix that they obtain similar but less significant results if
they use the continuous measure (in levels or logs) or an indicator for
any crop loss. The 200 TZS seems arbitrary and possibly ex post, since
it leads to the most attractive results of the three sets of results
provided. It would be useful to more transparently show readers the
underlying variation and possible non-linearities, perhaps with a binned
version of the variable.

[RESPONSE:]

\item The authors use causal language throughout the paper, describing
their results as effects. Of course, crop loss may be far from randomly
assigned, so this language is not appropriate.

[RESPONSE:]

[Matches with ADF's question on rainfall and with R2's first question 
on what drives crop losses]

Shamma: look for data on rainfall and papers that have used rainfall
for Kagera data

- problem is that we do not know when the crop loss occurred within
the survey period (which is 12 months for the first round and therefore
very long).

- there is no 30 year average to compare the observed district level
rainfall data to.

Crop loss question:

QB9: ``Have you lost any part of the harvested crop to insects,
rodents, fire, rotting, etc.?

\item In the regression specifications, the authors express the survey
wave fixed effects as $D_{i}'\alpha$, where $D_{i}$ is a vector of survey
wave dummies. The subscript is incorrect, and it would be much more
natural to just write $\alpha_{t}$.

[RESPONSE:] We have changed the equation and the associated text.

\item What other time-varying covariates are available in the survey? It
would be interesting to consider the extent to which the crop loss
results are explained by other household-level idiosyncratic variation.

[RESPONSE:] [Rainfall data analysis]

- no live stock shock question
- household member death
- price indices
- illness of mother and other household members

\item The authors report some nonsensical coefficients and standard
errors like -0.000 (0.000). All coefficients and standard errors should
be written to two significant digits.

[RESPONSE:] The 0.000 coefficients were predominantly for variables
involving initial assets. 
We have changed the variable definition for assets, so that it is
now measured in 1,000,000 TZS.

Note that there are still a few places---mainly for estimations
involving the use of modern contraceptives---where the estimated 
coefficients are so small that presenting additional digits does not make
much economic sense.
In those cases the standard errors are always equal to or larger than
the coefficients.
See, for example, Table A-8 on the effects of log crop loss.

\item Table 6 regresses pregnancy on contraceptive use, crop loss, and
their interaction. As the authors note, this regression is quite
difficult to interpret, given the endogeneity of contraceptive use. I
think this table adds little to the paper.

[RESPONSE:] Although we understand the concern about endogeneity, 
we believe the table is worth keeping. 
In the demography literature there is a substantial disagreement on whether
sustained fertility reduction requires access to modern contraceptives.
Part of the problem in this literature is that the ``intensity'' with
which a couple uses contraceptives---either traditional or modern---is
difficult to measure correctly.
[something on the panel nature of the data and that that is important
for getting at the incentives and the use over time; not perfect but
at least is for a shorter period than other (cross-sectional) surveys]
The table is important because it provides us a way to get at this
intensity by showing that only couples that have a strong incentive to
prevent pregnancies appear to be successful users of contraceptives.
Furthermore, this shows that it is possible to postpone pregnancies with
just traditional contraceptives because traditional contraceptives account 
for the vast proportion of contraception used.

[also speaks to whether this is a conscious decision or not] 
The table also supports our conjecture that the estimated effects 
[might need a less causal word] 
of crop loss capture a conscious decision. 
If shocks lead to lower probability of pregnancy and birth because
of other factors than the decision to use contraceptive we should see
a negative and statistically significant coefficient for the crop loss
dummy by itself.
We do not.
Instead, what we get is a substantial negative coefficient for the interaction
between contraceptive use and crop loss and a coefficient for crop loss 
on its own that is positive and far from statistically significant.

\item Several appendix tables interact crop loss with initial assets in
Round 1 of the survey. This exercise is interesting and worthy of doing,
given that one might expect asset-poor households to have particular
difficulty in smoothing consumption. But the authors should leave Round
1 outcomes out of this regression.

[RESPONSE:] We have added the results where round 1 has been excluded
to the relevant tables.
The results are broadly consistent with what we originally found, 
although substantially less likely to be statistically significant 
because of the lower variation in crop loss and the smaller sample
size.

[Already ran these; might be worthwhile having the results as an extra
panel in the relevant tables]

\end{enumerate}

% Overall, this is a simple but interesting paper that sheds light on
% important questions in the economics of fertility decisions before
% widespread adoption of modern contraception. It will make an incremental
% but solid contribution to the literature.

\newpage

\section*{Reviewer 2}

% This manuscript explores the effects of "unexpected" crop loss on
% fertility issues. The data come from four survey waves conducted in
% Tanzania, where there are repeated observations for the same individual.
% There are three main outcomes: births, pregnancies, and contraceptive
% use. The main estimation strategy relies on family fixed effects to
% address fixed differences (in wealth, education, etc.) that could be
% related to propensity to experience a crop loss. The crucial identifying
% assumption is that the timing of crop loss for a family is random and
% not related to other factors that might affect fertility.

The manuscript has a big positive: Fertility issues in developing
countries likely play an important role in economic growth, especially
via total completed fertility. Even if the shock in question did not
affect total completed fertility (i.e. individuals only postpone having
children), this paper would still be important since timing of births
matter. For example, being born in a period of low "resources" (food,
income, parental time) could irreparably harm a child's long-term
potential. To my knowledge, we know very little about the main
determinants of birth timing in a developing country context.

I have laid out some comments below in the interest of helping the
authors improve on a promising manuscript.

COMMENTS

\begin{description}

\item Omitted variables

Even with family fixed effects, crop loss is potentially tied with
omitted variables. Crop loss is subjectively measured, so we'd need to
rely on the claim of the survey respondent that the loss was for
"uncontrollable" circumstances. The use of family fixed effects does
address fixed differences in wealth or education that might be related
to families at greater risk of crop loss. But, there could still be time
varying factors. We can't be certain that the crop loss is not the
result of some disease epidemic (or conflict) that also could affect
fertility.

Suggestion: The manuscript should be a little more forthcoming about
this potential problem. I would like to see an exploration into the
potential mechanisms behind the crop loss. Is it weather?

[RESPONSE:]

[see response to R1's q5 on causal (or something like that)]

Weather is likely an important factor, but unfortunately rainfall data
are only available at district level (there are five districts in the
data) and even that has missing information.

\item Trend

I am concerned by the fact that crop loss is much higher in the first
and second waves of the survey (see Table 1 on page 6). So, there's a
chance that the estimates are driven by some time trend that affects
family types that are more likely to experience crop losses. (The survey
wave fixed effects won't deal with this if there are differential trends
in the type of families who experience crop loss.)

Suggestion: Control for family characteristics (age, proxies for
education, proxies for wealth) at the first wave interacted with survey
wave. At the very least, this will help acknowledge the identification
threat for the reader.

[RESPONSE:]

Claus: run regression with time trend interacted with wave 1 household characteristics

[need something on weather here?]

It is, indeed, concerning that there are so substantial
differences in the proportion of households that experienced a 
shocks over the different rounds, although it is possible that
this simply is a stark reminder of the volatile nature of life
in a developing country when dependent on rain-fed agriculture.
We have added an additional table to the Appendix to address this
concern. In addition to your suggestion of interacting a linear
time trend with wave 1 characteristics (assets per capita,
age group dummies, and education group dummies), we also present
a separate model where wave dummies are interacted with the
same set of wave 1 characteristics.
The small differences in point estimate and level of significance 
indicate that our results are not driven by differential trends
in which types of households that experience crop loss.

[Add something on this in paper?]

\item Spillovers

Neighbors may be affected an individuals crop loss. For one, price of
food goes up. This could be a positive income shock for families with
crops, but bad for families working in other areas.

Suggestion: Aggregate crop loss to the community level and use that as
the treatment variable.

[RESPONSE:]

\item Births analysis

The births analysis automatically drops first wave of survey due to the
longer lag in reported crop loss (Table 3). However, I did not see this
mentioned anywhere.

[RESPONSE:]

\end{description}

\newpage
\bibliographystyle{econometrica}
\bibliography{../paper/incomeShocks-jde-r1}
\addcontentsline{toc}{section}{References}

\end{document}